---
title: "Policy Recommendations for the Republican Party"
author: "Dagmawi Zerihun, Dong Dong, Luke Landis"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  warning = F,
  message = F,
  dpi = 500,
  out.width = "90%"
)
```

## Introduction

In the rapidly evolving political landscape of the United States, the Republican Party faces a pivotal challenge: How can it effectively balance the diverse views within its ranks while striving for electoral success? This question lies at the heart of our analysis, commissioned by the Republican Party itself. As political consultants, our task is to dissect the party's internal dynamics, understand its core strengths, and identify potential fault lines that could impact its future strategies.To achieve this, we will leverage the rich dataset from the "What The Hell Happened?" project by Data for Progress (DFP). This project offers a detailed glimpse into the opinions and attitudes of party members, providing us with a comprehensive platform to analyze the Republican Party's current standing. Despite DFP's left-leaning inclination, their survey methodology is rigorous, making their data a valuable asset for our analysis.Our focus will be twofold: First, we will explore the issues that unite and divide the party. Then, we will examine the party's attitudes toward minority and underserved groups in the US. This dual approach will enable us to offer nuanced advice on which issues the party should prioritize and which to handle with more sensitivity.With the backdrop of the party's complex challenges outlined, we now turn our attention to the robust dataset from the 'What The Hell Happened?' project. This section will detail our methodological approach, setting the stage for a nuanced understanding of the party's current standing and potential pathways forward.

## Data and Design

The dataset comprises 3,215 interviews from registered voters, weighted to reflect the national voter demographics from 2018. The variables include measures of political participation, policy positions, and attitudes toward various demographic groups. We will measure partisan membership using the party identification variable provided in the dataset. The size of the Republican Party within the sample will be assessed by counting the number of respondents who identify as Republicans. The demographic profile of the party will be analyzed based on race, education, and age distributions within the Republican-identifying respondents. Armed with a comprehensive dataset reflective of national voter demographics, our analysis will now delve into the heart of the Republican Party. We begin by examining the overarching characteristics of the party, followed by a deeper exploration of the issues that unite and divide its members, and their perspectives on diverse demographic groups. 

## Analysis

### Overview of the Party

The visualizations below break down the demographic profile of the Republican party by race, education, and gender. From the race and gender graph, it appears that the majority of respondents who identify as Republicans are White, regardless of gender. There is a smaller representation of Black, Latino/a, Asian, Native American, Middle Eastern, Mixed Race, and Other categories among the Republican respondents.

As for education, the majority of Republican respondents across both genders have some college education or have obtained a high school diploma. There is a noticeable presence of those with a 2-year degree, 4-year degree, and postgraduate education as well, indicating a diverse range of educational backgrounds within the party.

```{r fig1, fig.cap="Fig. 1:Breakdown of Survey Respondents", fig.width=10, fig.height=6}
## data
library(tidyverse)
library(socsci)
url <- "http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv"
Data <- read_csv(url)
# Assuming Data is already loaded and in the correct format
# Reorder the factor levels for educ_cat
## recodes
Data <- Data |>
  mutate(
    ## education
    educ_cat = frcode(
      educ == 1 ~ "No HS",
      educ == 2 ~ "HS",
      educ == 3 ~ "Some College",
      educ == 4 ~ "2-year Degree",
      educ == 5 ~ "4-year Degree",
      educ == 6 ~ "Postgrad"
    ),
    ## race
    race_new = frcode(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Latino/a",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 8 ~ "Middle Eastern",
      race == 6 ~ "Mixed Race",
      TRUE ~ "Other"
    ),
    ## party id
    pid3_new = frcode(
      pid3 == 1 ~ "Democrat",
      !(pid3 %in% 1:2) ~ "Independent/Other",
      pid3 == 2 ~ "Republican"
    ),
    ## gender
    gender_id = ifelse(
      gender == 1, "Male", "Female"
    ),
    ## age
    age = 2018 - birthyr,
    ## issues
    across(
      c(ICE, BAIL_item, WELTEST, PUBLICINT, GREENJOB, POLFEE),
      ~ frcode(
        .x == 5 ~ "Strongly oppose",
        .x == 4 ~ "Oppose",
        .x == 3 ~ "Neither",
        .x == 2 ~ "Support",
        .x == 1 ~ "Strongly support"
      )
    )
  )
Data$educ_cat <- factor(Data$educ_cat, levels = c("No HS", "HS", "Some College", "2-year Degree", "4-year Degree", "Postgrad"))

Data %>%
  group_by(gender_id, pid3_new) %>%
  ct(educ_cat, cum = T) %>% # computes cumulative pct
  ggplot(aes(
    x = pct,
    y = pid3_new,
    fill = educ_cat
  )) +
  geom_col() +
  facet_wrap(~ gender_id) +
  scale_x_continuous(labels = percent) +
  scale_fill_manual(values = c("#DAE8FC", "#BDD7EE", "#84B0D6", "#5593C7", "#3272B8", "#204E8A")) +  # Maintain the original color order
  labs(
    x = NULL,
    y = NULL,
    title = "Breakdown of respondents by Education, Partisanship and Gender",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = NULL
  ) +
  theme_minimal()
```


```{r fig2, fig.cap=" Fig. 2:Breakdown of Survey Respondents", fig.width=10, fig.height=6}
Data |>
  group_by(gender_id, pid3_new) |>
  ct(race_new, cum = T) |> # computes cumulative pct
  ggplot() +
  aes(
    x = pct,
    y = pid3_new,
    fill = race_new
  ) +
  geom_col() +
  facet_wrap(~ gender_id) +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Breakdown of respondent race by partisanship\nand gender",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = NULL
  )+
  theme_minimal()
```

### Major Issues:
In the below visualizations , we use the survey responses to evaluate a wide array of policy concerns , grouping responses by broader policy categories.
The first visualization depicts public opinion on two immigration-related policy proposals. The first policy, "Deport Undocumented Immigrants," shows a significant portion of respondents indicating 'Strongly support,' followed by a slightly smaller, yet substantial, portion who 'Strongly oppose' the policy. The 'Somewhat support' category is the next largest, while fewer respondents are in the 'Somewhat oppose' and 'Neither support nor oppose' categories. The second policy, "De-fund ICE" (Immigration and Customs Enforcement), presents a contrasting distribution of opinions. The largest segment of respondents 'Strongly oppose' the policy, followed by a smaller proportion that 'Strongly support' it. The 'Somewhat oppose' category is slightly larger than 'Somewhat support.' A minimal portion of respondents indicate neutrality with 'Neither support nor oppose.'Overall, the chart demonstrates a divided public opinion on these immigration issues, with stronger feelings expressed either in support of deporting undocumented immigrants or in opposition to defunding ICE.

```{r fig3, fig.cap=" Fig. 3:Immigration Policies", fig.width=10, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)
# Read the data
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Recoding the 'pid7' variable for party identification
cces <- cces %>%
  mutate(pid_new = case_when(
    pid7 == 5 ~ "Lean Republican",
    pid7 == 6 ~ "Not very strong Republican",
    pid7 == 7 ~ "Strong Republican"
  ))

# Filter for Republican respondents and exclude NA values
republicans <- cces %>%
  filter(pid_new %in% c("Lean Republican", "Not very strong Republican", "Strong Republican")) %>%
  filter(!is.na(DEPORT), !is.na(ICE))

# Recode the variables for clarity
republicans <- republicans %>%
  mutate(
    deport_view = case_when(
      DEPORT == 1 ~ "Strongly support",
      DEPORT == 2 ~ "Somewhat support",
      DEPORT == 3 ~ "Neither support nor oppose",
      DEPORT == 4 ~ "Somewhat oppose",
      DEPORT == 5 ~ "Strongly oppose",
      DEPORT == 6 ~ "Don't know"
    ),
    ice_view = case_when(
      ICE == 1 ~ "Strongly support",
      ICE == 2 ~ "Somewhat support",
      ICE == 3 ~ "Neither support nor oppose",
      ICE == 4 ~ "Somewhat oppose",
      ICE == 5 ~ "Strongly oppose",
      ICE == 6 ~ "Not sure"
    ) 
  )%>%
  filter(deport_view != "Don't know", ice_view != "Not sure")
republicans|>
  pivot_longer(
    cols = c(ice_view, deport_view)
  ) |>
  mutate(
    name = frcode(
      name == "ice_view" ~ "De-fund ICE",
      name == "deport_view" ~ "Deport Undocumented Immigrants"
    )
  ) |>
  group_by(name) |>
  ct(value, show_na = F, cum = T) |>
  ggplot() +
  aes(x = pct, y = name, fill = value) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Support for different immigration related policy proposals",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = NULL
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = T)
  )
```

The provided chart delineates the perspectives of Republican respondents on gun control, segmented by age. It shows a spectrum of opinions ranging from support for the current regulations to various degrees of advocating for more or less stringent laws. The older demographic, particularly those aged 65 and above, predominantly feel that the current regulations are appropriate. In contrast, the youngest cohort, aged 18-29, shows a notable inclination towards making it more difficult to purchase all types of guns. The middle age groups (30-39, 40-49, and 50-64) display mixed views that span across making it more difficult to buy some or all types of guns, to making it less difficult to buy some types. The chart reveals an evident generational divide in the attitudes towards gun control within the Republican respondents.

```{r fig4, fig.cap=" Fig. 4:Gun Regulations", fig.width=10, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)  # For percentage formatting

# Read the data
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Recoding the 'pid7' variable for party identification
cces <- cces %>% 
  mutate(pid_new = case_when(
    pid7 == 5 ~ "Lean Republican",
    pid7 == 6 ~ "Not very strong Republican",
    pid7 == 7 ~ "Strong Republican"
  ))

# Filter for Republican respondents and exclude NA values in GUNS
republicans <- cces %>% 
  filter(pid_new %in% c("Lean Republican", "Not very strong Republican", "Strong Republican"),
         !is.na(GUNS))

# Recode the GUNS variable for clarity
republicans <- republicans %>% 
  mutate(guns_view = case_when(
    GUNS == 1 ~ "More difficult to buy all types",
    GUNS == 2 ~ "More difficult to buy some types",
    GUNS == 3 ~ "Current regulations are right",
    GUNS == 4 ~ "Less difficult to buy some types",
    GUNS == 5 ~ "Less difficult to buy all types",
    GUNS == 6 ~ "Not sure"
  ))

# Recode the age groups
republicans <- republicans %>%
  mutate(age_group = case_when(
    age5 == 1 ~ "18-29",
    age5 == 2 ~ "30-39",
    age5 == 3 ~ "40-49",
    age5 == 4 ~ "50-64",
    age5 == 5 ~ "65+"
  ))

# Calculating percentages instead of counts
republicans_grouped <- republicans %>% 
  count(guns_view, age_group) %>% 
  group_by(age_group) %>% 
  mutate(perc = n / sum(n) * 100)

# Create a grouped bar chart for gun regulation views by age group with percentages
ggplot(republicans_grouped, aes(x = guns_view, y = perc, fill = age_group)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +  # Using a color palette for distinction
  coord_flip() +  # Flips the coordinates for better readability
  labs(title = "Views on Gun Regulations Among Republican Respondents by
       Age Group (Percentages)",
       x = "View on Gun Regulations",
       y = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.title = element_blank()) +  # Improve label readability and adjust legend
  scale_y_continuous(labels = function(x) paste0(x, "%"))  # Correctly format y-axis as percentages
```

The below visualization presents public support for two health care-related policies: the production of life-saving drugs and Medicare. The response categories include 'Support', 'Strongly support', 'Strongly oppose', 'Oppose', and 'Neither' (neither support nor oppose).For the production of life-saving drugs, there is a notable majority that either supports or strongly supports the policy, with 'Strongly support' being the most prevalent stance. There is a smaller but significant proportion of respondents who 'Strongly oppose' or 'Oppose' the policy, and a very small percentage are neutral.Regarding Medicare, the pattern of support is similar, with the majority of respondents either supporting or strongly supporting the policy. 'Support' appears to be the predominant response, followed closely by 'Strongly support'. Opposition to Medicare is less than that for the production of life-saving drugs, and the proportion of respondents who are neutral is similarly small.The visualization underscores a strong backing for both health care-related policies, with a greater consensus on the production of life-saving drugs compared to Medicare.

```{fig5, fig.cap=" Fig. 5:Healthcare Policies", fig.width=10, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)



library(tidyverse)
library(socsci)
## add any other packages, like coolorrr, you want to use

## read in your data here:
url <- "http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv"
Data <- read_csv(url)
## do your data cleaning and recoding here: 
# recodes
Data <- Data |>
  mutate(
    ## education
    educ_cat = frcode(
      educ == 1 ~ "No HS",
      educ == 2 ~ "HS",
      educ == 3 ~ "Some College",
      educ == 4 ~ "2-year Degree",
      educ == 5 ~ "4-year Degree",
      educ == 6 ~ "Postgrad"
    ),
    ## race
    race_new = frcode(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Latino/a",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 8 ~ "Middle Eastern",
      race == 6 ~ "Mixed Race",
      TRUE ~ "Other"
    ),
    ## party id
    pid3_new = frcode(
      pid3 == 1 ~ "Democrat",
      !(pid3 %in% 1:2) ~ "Independent/Other",
      pid3 == 2 ~ "Republican"
    ),
    ## gender
    gender_id = ifelse(
      gender == 1, "Male", "Female"
    ),
    ## age
    age = 2018 - birthyr,
    ## issues
    across(
      c(ICE, BAIL_item, WELTEST, PUBLICINT, GREENJOB, POLFEE),
      ~ frcode(
        .x == 5 ~ "Strongly oppose",
        .x == 4 ~ "Oppose",
        .x == 3 ~ "Neither",
        .x == 2 ~ "Support",
        .x == 1 ~ "Strongly support"
      )
    )
  )
Data_Filter  <-Data |> filter( pid3_new == 'Republican')


# Adjust your data processing here accordingly
Data_Filter %>%
  mutate(across(
    c(M4A, PUBLICGEN),
    ~ case_when(
      .x == 5 ~ "Strongly oppose",
      .x == 4 ~ "Oppose",
      .x == 3 ~ "Neither",
      .x == 2 ~ "Support",
      .x == 1 ~ "Strongly support"
    )
  )) %>%
  pivot_longer(
    cols = c(M4A, PUBLICGEN)
  ) %>%
  mutate(
    name = case_when(
      name == "M4A" ~ "Medicare",
      name == "PUBLICGEN" ~ "Production of life-saving drugs"
    )
  ) %>%
  group_by(name) %>%
  ct(value, show_na = FALSE, cum = TRUE) %>%
  ggplot() +
  aes(x = pct, y = name, fill = value) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Support for health care related policies",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = NULL
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = T)
  )
```

The visualizations present data on the level of support for Donald Trump among Republican age groups and their inclination to re-elect him. The first chart, titled "Support for Trump Among Republicans by Age Group," shows varying degrees of support categorized as "strongly" and "somewhat" for both Trump and the Republican Party, along with a "Not sure" category, across five age groups: 18-29, 30-39, 40-49, 50-64, and 65+. The second chart, titled "Would Republicans Re-elect Trump?", depicts the percentage of Republicans in the same age groups who would vote for Trump if given the chance. The vsualizations suggest that older age groups tend to show stronger support for Trump and a higher likelihood of re-electing him, compared to the younger demographics.

```{r fig6, fig.cap=" Fig. 6: Republicans on Supporting Donald Trump"  , fig.width=10, fig.height=6}
library(tidyverse)
library(scales)  # For percentage formatting
library(RColorBrewer)  # For color palettes

# Read the data

Data <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Recoding the 'pid3' variable for party identification
Data <- Data %>% 
  mutate(
    pid3_new = case_when(
      pid3 == 1 ~ "Democrat",
      pid3 == 2 ~ "Republican",
      TRUE ~ "Independent/Other"
    ),
    # Assuming 'age' is already present in the dataset
    # Recode the age groups
  age_group = case_when(
    age5 == 1 ~ "18-29",
    age5 == 2 ~ "30-39",
    age5 == 3 ~ "40-49",
    age5 == 4 ~ "50-64",
    age5 == 5 ~ "65+"
  ),
    # Recoding support for Trump vs Republican Party
    dtrpsupp_recode = case_when(
      dtrpsupp == 1 ~ "Trump supporter, strongly",
      dtrpsupp == 2 ~ "Trump supporter, somewhat",
      dtrpsupp == 3 ~ "Republican Party supporter, somewhat",
      dtrpsupp == 4 ~ "Republican Party supporter, strongly",
      dtrpsupp == 5 ~ "Not sure"
    ),
    djtrelct_recode = case_when(
      djtrelct == 1 ~ "Yes",
      djtrelct == 2 ~ "No",
      djtrelct == 3 ~ "Not sure"
    )
  )

# Filter for Republican respondents and exclude NA values in 'dtrpsupp_recode'
republicans <- Data %>% 
  filter(pid3_new == "Republican", !is.na(dtrpsupp_recode))
# Assuming Data is already loaded and filtered for Republicans with necessary recodes
# Calculating percentages for 'dtrpsupp_recode' instead of counts
republicans_grouped <- Data %>%
  filter(pid3_new == "Republican", !is.na(dtrpsupp_recode)) %>%
  count(dtrpsupp_recode, age_group) %>%
  group_by(age_group) %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() # Remove the grouping


# Create a bar chart for support for Trump vs Republican Party by age group with percentages
ggplot(republicans_grouped, aes(x = dtrpsupp_recode, y = perc, fill = age_group)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +  # Automatically convert to percentage scale
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Support for Trump Among Republicans by Age Group",
       x = "Type of Support",
       y = "Percentage") +
  coord_flip() +  # Flips the coordinates for better readability
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r fig=7, fig.cap=" Fig. 7:Trump Re-election", fig.width=10, fig.height=6}
# Filter for Republican respondents and exclude NA values in 'dtrpsupp_recode'
republicans <- Data %>% 
  filter(pid3_new == "Republican", !is.na(djtrelct_recode))
# Assuming Data is already loaded and filtered for Republicans with necessary recodes
# Calculating percentages for 'dtrpsupp_recode' instead of counts
republicans_grouped <- Data %>%
  filter(pid3_new == "Republican", !is.na(djtrelct_recode)) %>%
  count(djtrelct_recode, age_group) %>%
  group_by(age_group) %>%
  mutate(perc = n / sum(n)) %>%
  ungroup() # Remove the grouping

# Create a bar chart for support for Trump vs Republican Party by age group with percentages
ggplot(republicans_grouped, aes(x = djtrelct_recode, y = perc, fill = age_group)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +  # Automatically convert to percentage scale
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Would Republicans Re-elect Trump?",
       x = "Type of Support",
       y = "Percentage") +
  coord_flip() +  # Flips the coordinates for better readability
  theme_minimal() +
  theme(legend.position = "bottom")
```

The below visualization displays public opinion on two policies: pollution fees on high-emission companies and jobs in energy-efficient infrastructure.The bar graph indicates that a significant majority strongly supports both policies, with substantial support for pollution fees and a slightly lesser, yet still majority, support for jobs in energy-efficient infrastructure.

```{r fig8, fig.cap=" Fig. 8: Environmental and Economic Policies", fig.width= 12, fig.height=6, out.width='100%', out.height= '100%'}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)

# Read your data here
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Adjust your data processing here accordingly
cces %>%
  filter(GREENJOB != 6, POLFEE != 6) %>%
  mutate(across(
    c(GREENJOB, POLFEE),
    ~ case_when(
      .x == 5 ~ "Strongly oppose",
      .x == 4 ~ "Somewhat oppose",
      .x == 3 ~ "Neither support nor oppose",
      .x == 2 ~ "Somewhat support",
      .x == 1 ~ "Strongly support"
    )
  )) %>%
  pivot_longer(
    cols = c(GREENJOB, POLFEE)
  ) %>%
  mutate(
    name = case_when(
      name == "GREENJOB" ~ "Job in energy-efficient infrastructure",
      name == "POLFEE" ~ "Pollution fees on high-emission companies",
    )
  ) %>%
  group_by(name) %>%
  ct(value, show_na = FALSE, cum = TRUE) %>%
  ggplot() +
  aes(x = pct, y = name, fill = value) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Support for Environmental and Economic Policies",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = "Response"
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = T),
  )
  ggsave("plot.png", width=12, height=6, dpi=300)
```

The below visualization assesses the level of support for three different policies related to marijuana: release for marijuana-related offenses, reinvestment of marijuana tax revenues, and national legalization of marijuana. The pattern of responses is similar across all three policies, with a majority expressing strong support, a moderate level showing somewhat support, and smaller fractions indicating opposition.

```{r fig9, fig.cap=" Fig. 9: Marijuana", fig.width=12, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)

# Read your data here
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Adjust your data processing here accordingly
cces %>%
  filter(!is.na(MARREP), !is.na(MARAM), !is.na(MARLEG)) %>%
  mutate(across(
    c(MARREP, MARAM, MARLEG),
    ~ case_when(
      .x == 5 ~ "Strongly oppose",
      .x == 4 ~ "Somewhat oppose",
      .x == 3 ~ "Neither support nor oppose",
      .x == 2 ~ "Somewhat support",
      .x == 1 ~ "Strongly support"
    )
  )) %>%
  pivot_longer(
    cols = c(MARREP, MARAM, MARLEG)
  ) %>%
  mutate(
    name = case_when(
      name == "MARREP" ~ "Reinvestment of marijuana tax revenues",
      name == "MARAM" ~ "Release for marijuana-related offenses",
      name == "MARLEG" ~ "National legalization of marijuana"
    )
  ) %>%
  group_by(name) %>%
  ct(value, show_na = FALSE, cum = TRUE) %>%
  ggplot() +
  aes(x = pct, y = name, fill = value) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Public Opinion on Marijuana Policies",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = "Response"
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = T)
  )

```

The below visualization "Republicans' on Economic and Social Policies Related to Wealth Distribution through Taxation," illustrates the stance of Republicans on four different economic and social policies. These include a tax on wealth greater than USD 100 million, government production of life-saving drugs, free college funded by tax on high incomes, and a USD 5,000 savings account at birth. The graph suggests a polarization of opinions with significant strong support and strong opposition for the wealth tax and free college policies, while the support appears more balanced for the production of life-saving drugs and the savings account at birth initiative.

```{r fig10, fig.cap=" Fig. 10:Taxation of the Wealthiest", fig.width=12, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)

# Read your data here
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Adjust your data processing here accordingly
cces %>%
  filter(!is.na(PUBLICGEN), !is.na(BOND), !is.na(FREECOLL), !is.na(WEALTH)) %>%
  mutate(across(
    c(PUBLICGEN, BOND, FREECOLL, WEALTH),
    ~ case_when(
      .x == 5 ~ "Strongly oppose",
      .x == 4 ~ "Somewhat oppose",
      .x == 3 ~ "Neither support nor oppose",
      .x == 2 ~ "Somewhat support",
      .x == 1 ~ "Strongly support"
    )
  )) %>%
  pivot_longer(
    cols = c(PUBLICGEN, BOND, FREECOLL, WEALTH)
  ) %>%
  mutate(
    name = case_when(
      name == "PUBLICGEN" ~ "Government production of life-saving drugs",
      name == "BOND" ~ "$5,000 savings account at birth",
      name == "FREECOLL" ~ "Free college funded by tax on high incomes",
      name == "WEALTH" ~ "Tax on wealth greater than $100 million"
    )
  ) %>%
  group_by(name) %>%
  ct(value, show_na = FALSE, cum = TRUE) %>%
  ggplot() +
  aes(x = pct, y = name, fill = value) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Republicans' on Economic and Social Policies 
    Related to Wealth Distribution through Taxation",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = "Response"
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = T)
  )

```

In the below visualization, "Republican Respondents' Opinion on Various Policies," the opinions of Republican respondents on four policies are showcased: ending US involvement in Yemen, ending solitary confinement, drug test for government aid, and automatic voter registration. The graph reveals a diversity of opinions, with notable strong support for ending involvement in Yemen and strong opposition to automatic voter registration, indicating a split in views on these policies among Republicans.

```{r fig11, fig.cap=" Fig. 11:Other Policies", fig.width=12, fig.height=6}
# Load necessary libraries
library(tidyverse)
library(scales)
library(ggplot2)
library(dplyr)
library(tidyr)
library(socsci)
library(coolorrr)

## data viz settings
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)

# Read your data here
cces <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

# Adjust your data processing here accordingly
cces %>%
  filter(AVR != 6, YEMEN != 6, SOLITARY != 6, WELTEST != 6) %>%
  mutate(across(
    c(AVR, YEMEN, SOLITARY, WELTEST),
    ~ case_when(
      .x == 5 ~ "Strongly oppose",
      .x == 4 ~ "Somewhat oppose",
      .x == 3 ~ "Neither support nor oppose",
      .x == 2 ~ "Somewhat support",
      .x == 1 ~ "Strongly support"
    )
  )) %>%
  pivot_longer(
    cols = c(AVR, YEMEN, SOLITARY, WELTEST),
    names_to = "Policy",
    values_to = "Support_Level"
  ) %>%
  mutate(
    Policy = case_when(
      Policy == "AVR" ~ "Automatic Voter Registration",
      Policy == "YEMEN" ~ "Ending US involvement in Yemen",
      Policy == "SOLITARY" ~ "Ending Solitary Confinement",
      Policy == "WELTEST" ~ "Drug Test for Government Aid"
    )
  ) %>%
  group_by(Policy) %>%
  count(Support_Level) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot() +
  aes(x = pct, y = Policy, fill = Support_Level) +
  geom_col() +
  scale_x_continuous(
    labels = scales::percent
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Republican Respondents' Opinion on Various Policies",
    caption = "Source: What the Hell Happened? (Data for Progress)",
    fill = "Support Level"
  ) +
  ggpal(
    aes = "fill",
    guide = guide_legend(reverse = TRUE)
  )

```

### Views on Underrepresented Groups
Having identified the core issues that garner agreements and differences within the Republican party, our focus shifts to another critical dimension: the party's stance towards minority and underserved groups.
The visualization titled "Republican attitudes toward minority or underserved groups in the US" depicts three histograms, each corresponding to a different sentiment: Fear of Demographic Change, Hostile Sexism, and Racial Animus. The histograms display the frequency of responses on a scale from 0 to 1, where 0 likely indicates no fear or hostility, and 1 indicates extreme fear or hostility. The Fear of Demographic Change histogram shows a prominent peak around the 0.75 mark, suggesting a concentration of higher fear levels among respondents. Hostile Sexism displays a more spread out distribution with a notable peak just below 0.75, indicating a higher level of sexist attitudes. The Racial Animus histogram also shows a spread but with significant counts in the middle of the scale, suggesting relatively lower levels of racial animosity among respondents. Collectively, these histograms indicate that while there is a degree of unity in lower levels of fear regarding demographic change, there is more division in attitudes toward sexism and racial animus, with responses ranging widely across the spectrum. As these scaled variables are aggregates of many differnt questions regarding underserved groups in the survey, the higher scaled values in two of the three variables indicates that the views on underrepresented groups in the US among republicans is not too positive.

```{r fig12, fig.cap=" Fig. 12: Attitude Towards Underserved Groups", fig.width=12, fig.height=6}
library(ggrain)
# Assuming your data frame is named 'Data' and you have 'party_affiliation' and 'age' variables.
Data <- read.csv("http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv")

library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)

library(tidyverse)
library(socsci)
## add any other packages, like coolorrr, you want to use

## read in your data here:
url <- "http://filesforprogress.org/datasets/wthh/DFP_WTHH_release.csv"
Data <- read_csv(url)
## do your data cleaning and recoding here: 
# recodes
Data <- Data |>
  mutate(
    ## education
    educ_cat = frcode(
      educ == 1 ~ "No HS",
      educ == 2 ~ "HS",
      educ == 3 ~ "Some College",
      educ == 4 ~ "2-year Degree",
      educ == 5 ~ "4-year Degree",
      educ == 6 ~ "Postgrad"
    ),
    ## race
    race_new = frcode(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Latino/a",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 8 ~ "Middle Eastern",
      race == 6 ~ "Mixed Race",
      TRUE ~ "Other"
    ),
    ## party id
    pid3_new = frcode(
      pid3 == 1 ~ "Democrat",
      !(pid3 %in% 1:2) ~ "Independent/Other",
      pid3 == 2 ~ "Republican"
    ),
    ## gender
    gender_id = ifelse(
      gender == 1, "Male", "Female"
    ),
    ## age
    age = 2018 - birthyr,
    ## issues
    across(
      c(ICE, BAIL_item, WELTEST, PUBLICINT, GREENJOB, POLFEE),
      ~ frcode(
        .x == 5 ~ "Strongly oppose",
        .x == 4 ~ "Oppose",
        .x == 3 ~ "Neither",
        .x == 2 ~ "Support",
        .x == 1 ~ "Strongly support"
      )
    )
  )
set_theme()   
set_palette(
  qualitative = c("steelblue", "gray", "red3", "purple4", "gold2"),
  sequential = c("white", "steelblue"),
  diverging = c("red3", "white", "steelblue"),
  binary = c("steelblue", "red3"),
  from_coolors = F
)

# Assuming data is already loaded into a dataframe named 'data
Data_Filter <- Data |> filter( pid3_new == 'Republican')
result <- Data_Filter   %>% 
  select(racial_resentment_scaled, hostile_sexism_scaled, fear_of_demographic_change_scaled ) %>%
  pivot_longer(
    cols = racial_resentment_scaled:fear_of_demographic_change_scaled 
  )
ggplot(result, aes(x = value)) +
  geom_histogram(binwidth = 0.1, position = "identity", alpha = 0.7, color = "black") +
  facet_grid(~ name, scales = "free", labeller = labeller(name = c("hostile_sexism_scaled" = "Hostile Sexism", "racial_resentment_scaled" = "Racial Animus","fear_of_demographic_change_scaled" = "Fear Of Demographic Change"))) +
  labs(
   x = NULL,
    y = NULL,
    title = " Republican attitudes toward minority or underserved groups in the US",
    caption = "Source: What the Hell Happened? (Data for Progress)"
  )+ 
  theme_minimal()
```


## Conclusion
Through our analysis, it's clear that the Republican Party stands at a crossroads. The data-driven insights gleaned from the "What The Hell Happened?" project have illuminated key aspects of the party's internal dynamics and public perception. Our findings suggest that while there are issues of consensus that can serve as unifying forces, there are also significant challenges in addressing the concerns and perspectives of underrepresented groups.To navigate these complexities, we recommend a balanced approach. The party should leverage its consensus on major issues to strengthen its core base, while simultaneously engaging in open dialogues and inclusive policies that resonate with minority and undeserved communities. This dual strategy will not only reinforce the party's traditional strengths but also broaden its appeal to a more diverse electorate, ensuring its relevance and vitality in the evolving political landscape.The Republican Party's path forward requires a thoughtful blend of tradition and adaptation. By embracing both the voices within and the changing demographics outside, the party can chart a course towards sustained success and relevance in the American political arena.

## Works Cited
https://github.com/milesdwilliams15/Teaching/blob/main/DPR%20101/Notes/15_survey_data_pt2.md